---
title: "CSCI 3701 Project Step 3"
author: "Paul Friederichsen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, error=FALSE)
library(tidyverse)
library(rlist)
library(plotly)

library(igraph)
library(visNetwork)


chunk2 <- function(x,n) split(x, cut(seq_along(x), n, labels = FALSE)) 
splitdf <- function(df, n) split(df, (seq(nrow(df))-1) %/% n)
```

[GitHub Repo](https://github.com/floogulinc/csci-3701-project)

## Attempting to use Neo4j

My first step to working with Neo4j is to get my data into it.

### More data manipulation

Import the data from previous steps:
```{r}
site.data <- readRDS("sitedata.Rda")
edges.studentOf <- readRDS("edges-student-of.Rda")
edges.advisorTo <- readRDS("edges-advisor-to.Rda")
edges <- readRDS("alledges.Rda")
```

I would like to make a new dataframe for `nodes` and clean up whats included. I will also add a `country` field that is the first country of the first school of the person. We will assume that is their country.

```{r}
nodes <- site.data[c("id", "name")]
nodes$country <- map(site.data$schools, ~ .) %>% map(~ .$countries) %>% map(first) %>% map(first) %>% as.character()
nodes$country[nodes$country == "NULL"] <- NA
```


#### Making CSV files

To get the graph into Neo4j, `csv` files seem to be the standard way.

```{R, eval=FALSE}
write_csv(nodes, "nodes.csv")
write_csv(edges.studentOf, "edges-student-of.csv")
write_csv(edges.advisorTo, "edges-advisor-to.csv")
```

### Connecting to Neo4j

```{r, eval=FALSE}
library(neo4r)
con <- neo4j_api$new(
  url = "http://localhost:7474",
  user = "neo4j", 
  password = "<password>"
)
con$get_version()
```

### Trying to get data into Neo4j

You can't load CSVs like this...

```{r, eval=FALSE}
on_load_query <- 'CREATE (n:Person) SET n = row;'
load_csv(url = "nodes.csv", 
         con = con, header = TRUE, periodic_commit = 50, 
         as = "row", on_load = on_load_query)
```

This is the first method I tried for adding nodes, but it's really slow and I think there is better.

```{r, eval=FALSE}

nodecyphers <- apply(nodes, 1, function(x) {vec_to_cypher(x, "Person")})
for(cypher in nodecyphers) {
  call_neo4j(paste("CREATE", cypher), con)
}

```

This didn't end up working either

```{r, eval=FALSE}

#site.data$country <- map(site.data$schools, ~ .) %>% map(~ .$countries) %>% map(first) %>% map(first) %>% as.character()
#site.data$country[site.data$country == "NULL"] <- NA

nodes$name <- gsub("'", " ", nodes$name) #escape '

nodecyphers <- apply(nodes, 1, function(x) {paste("CREATE",vec_to_cypher(x, "Person"))})
writeLines(nodecyphers, "nodes.cypher")
send_cypher("nodes.cypher", con)

```

A really hard to find error that came up with this is people with `'` in their names breaks the syntax of the query. There were more errors like this.

Neo4j ended up being too difficult to work with for me and I gave up on using it here.


## Making a GraphML file

Exporting the graph as a GraphML file allows it to be used in some other programs that deal with graphs.

Lets make a new graph, with only `advisorTo` edges and the new nodes structure.

```{r}
net <- graph_from_data_frame(edges.advisorTo, vertices = nodes, directed = TRUE)
net
```

Now we can export the graph in GraphML format:
```{r, eval=FALSE}
write_graph(net, "advisor-to-graph.graphml", format="graphml")
```

GraphML can be imported by various programs, like [Gephi](https://gephi.org/) (which is very broken on my machine).

## More graph analysis and visualization in R

My plan now is to do some more in depth analysis of the graph I have made in R.

```{r}
edge_density(net)
```
Looking at edge density, the ratio of number of edges and number of possible edges, we find it is very small, which we should probably expect with this data.

### Cliques

Let's see how many max cliques there are.
```{r}
count_max_cliques(net)
```

Here I am looking at the largest cliques in the graph
```{r}
clique_num(net)

lq <- largest_cliques(net)
length(lq)
```
We see that the largest cliques formed are of 4 nodes and there are 17 of them. In other words, there are 17 groups where every person has some connection to the other 3.

Let's graph them.
```{r, out.width="100%"}
induced.subgraph(net, unlist(lq)) %>% visIgraph()
```



